---
title: "Sentiment Indicator and Sector Performance"
output: html_document
---

```{r setup, include=FALSE}

```

This report documents analytical steps used in the Paper presented in Zlin, 2017. 

Document is structured as follows:

- market data analysis
- 

Data from Blooomberg database was downloaded separately and are stored in *.txt* files. Import and cleaning is not described here, but can be found in sourced file *sentimentDataClean.R*.
```{r import1, warning=FALSE, message=FALSE}
source("sentimentDataClean.R") # import and cleaning, Bloomberg Data
data <- sentimentDataClean()
str(data)
```
Monthly data cover period from  `r min(data[["sentData"]]$Date)` to `r max(data[["sentData"]]$Date)`

By following code data from [yahoo](https://finance.yahoo.com/) are fetched. Only starting period is set. **TODO**: for final version *to* should be set.
```{r, importFinData, cache=TRUE, warning=FALSE, message=FALSE}
library(tidyquant)

vwData  <- tq_get("VOW.DE", get = "stock.prices", from = " 2005-01-01")
bmwData  <- tq_get("BMW.DE", get = "stock.prices", from = " 2005-01-01")
daiData  <- tq_get("DAI.DE", get = "stock.prices", from = " 2005-01-01")
```

We are interested in Adjusted-Close prices as they reflect dividend payments and share splits.

```{r, finDatamunging, warning=FALSE, message=FALSE}
vwClose <- vwData %>% select(date, vw=adjusted) %>% mutate(date = ymd(date) )
bmwClose <- bmwData %>% select(date, bmw=adjusted) %>% mutate(date = ymd(date) )
daiClose <- daiData %>% select(date, dai=adjusted) %>% mutate(date = ymd(date) )
manufData <- vwClose %>% 
 left_join(bmwClose, by="date") %>%
 left_join(daiClose, by="date")
```

```{r}
manufLong <- manufData %>% gather(., manuf, adjClose, -date)

manufLong %>%
 group_by(manuf, year=year(date), month=month(date)) %>%
 summarise(
  nobs = n(),
  minPrice = min(adjClose),
  maxPrice = max(adjClose),
  diff = maxPrice-minPrice,
  volat = sd(adjClose) 
)
```


## Time Series Visualisation
First time-series is in levels. We need to reshape the data to meet ggplot standards.

```{r, fig.width=10, fig.heigh=4}
manufLong <- manufData %>% gather(., manuf, adjClose, -date)

ggplot(manufLong, aes(x=date, y=adjClose, col=manuf) ) + 
 geom_line() + 
 facet_grid(.~manuf) + theme_bw()
```

Time series adjusted to initial date:

```{r, fig.width=10, fig.heigh=5}
manufDataScale <- manufData %>% 
 mutate(
  vw = vw/first(vw),
  bmw = bmw/first(bmw),
  dai = dai/first(dai)
 )

manufDataScaleLong <- manufDataScale %>% gather(., manuf, adjClose, -date)

manufDataScaleLong$manuf <- factor(
 manufDataScaleLong$manuf, levels=c("vw", "bmw", "dai"))

ggxbreaks <- as.Date(paste0(2015:2017, "/1/1"), format="%Y/%m/%d" )

ggplot(manufDataScaleLong, aes(x=date, y=adjClose, col=manuf) ) + 
 geom_line() +  theme_bw() + 
 scale_colour_discrete("Manufacturer", labels=c( "VW", "BMW", "Daimler")) + 
 scale_y_continuous("Adjusted price, normed") + 
 scale_x_date("Date", date_labels = "%Y", date_minor_breaks = "1 year")

```






