---
title: "Sentiment Indicator and Sector Performance"
output: html_document
---

This report documents analytical steps used in the Paper presented in Zlin, 2017. 

The premis of the paper is:

 > Sentiment indicators have predictive power on fluctuations in prices of major economic subject in the economy.

 > Volatility of prices responds differently when change occurs in low levels.

- As a sentiment indicators ZEW and IFO indexes were used. 
- On a company level three companies were selected. Composit index of selected on a macro level. **TODO** sectors level indicator, too? 

Document is structured as follows:

- Data import and assumptions
- Market data analysis, returns
- Sentiment analysis
- Forecasting model description and performance.

# Data Import and Assumptions

Two input data files were used. Both were obtained through the Bloomberg database. Although there are tools for receiving market data by R packages, Bloomberg was used (see *paperFinData.R* for analysis and description of problems, such as missing values). Data containing sentiment indexes and market perfomence indicators (prices) are stored separately in *.txt* files. Import and cleaning is not described here, but can be found in sourced file *sentimentDataClean.R*.

## Sentiment Data Indexes
**Sentiment Indexes ** 

```{r import1, warning=FALSE, message=FALSE}
source("sentimentDataClean.R") # import and cleaning, Bloomberg Data
data <- sentimentDataClean()
str(data)
```
Monthly data cover period from  `r min(data[["sentData"]]$Date)` to `r max(data[["sentData"]]$Date)`

## Performance of Economic Indicators
As the project is oriented on Automotive industry, Volkswagen, BMW and Daimler companies were selected. Analysis is accompanied by tracking composite DAX index. 

```{r importPerf, warning=FALSE, message=FALSE}

file <- "D:/UTB/MyPaper/SentimentPerformancePaper/Data/perfmormanceManufEconomy.txt"
perfDataIn <- read_tsv(file) %>%
  mutate(Date = dmy(Date))

perfData <- perfDataIn[complete.cases(perfDataIn), ] #removes Christmas value
```
From the following Figure can be observed problematic development of Volkswagen (VW) prices. These are two-fold:
 1. Speculative pricing during the perdion 8/2008 when VW acquired Porsche. 
 2. Sharp decline in 2015 after Diesel Gate. Both of these events are rather company-related and cannot be atributed to general economic environment sentiment. This concerns supply/demand attributes of the market rather than speculative nature of the market. TODO: Justify portfolio composition, remove VW from the company-level analysis and continue only with BMW and Daimler (DAI)?
 
To solve the first problem, unfaithful value traded during `r perfData[ which.max(perfData$VOW), "Date"]`, averege value of day before and after replaced this value. 

```{r}
perfData[which.max(perfData$VOW),"VOW"] <- 
  mean(perfData[(-1:1)+which.max(perfData$VOW),"VOW"]$VOW[-2])
```
# Market Data Analysis, Returns
In the first step, table (only first values showed here) summarising monthly min, max and differences in stock prices were computed. As the assessment of volatility, standard deviation was computed. This is not correct for the future analysis because of a non-stationarity of time series.


```{r}
manufLong <- perfData %>% gather(., manuf, adjClose, -Date)

manufLong %>%
  group_by(manuf, year=year(Date), month=month(Date)) %>%
  summarise(
    nobs = n(),
    minPrice = min(adjClose),
    maxPrice = max(adjClose),
    diff = maxPrice-minPrice,
    volat = sd(adjClose) 
  )

```

Visualisation TODO TODO TODO:
```{r}
manufLong %>% filter(manuf!="CDAX") %>%
ggplot(., aes(x=Date, y=adjClose, col=manuf) ) + 
  geom_line() + 
  facet_grid(.~manuf) + theme_bw()
```

Because of different values (levels) comparison might be misleading. So the scaling that for all times series value of share was set to 1 at 1/1/2015 = 1,

```{r}
perfDataScale <- perfData %>% 
  arrange(Date) %>%
  mutate(
    VW = VOW/first(VOW),
    BMW = BMW/first(BMW),
    DAI = DAI/first(DAI),
    DAX = CDAX/first(CDAX)
  ) %>% 
  select(Date, VW, BMW, DAI, DAX)

perfDataScale
```

Visualisation TOTOTOTOTOTO

```{r}
# ggdata prep
perfDataScaleLong <- perfDataScale %>% gather(., Title, adjClose, -Date)
# settings for plot
ggxbreaks <- as.Date(paste0(2015:2017, "/1/1"), format="%Y/%m/%d" )
# relevel to match top-down Time-Series
perfDataScaleLong$Title <- factor(
  perfDataScaleLong$Title, levels=c("VW", "DAX", "BMW", "DAI"))

ggplot(perfDataScaleLong, aes(x=Date, y=adjClose, col=Title) ) + 
  geom_line() + theme_bw()+ 
  scale_colour_discrete("Title", labels=c("VW", "DAX", "BMW", "Daimler")) + 
  scale_y_continuous("Adjusted price, normed") + 
  scale_x_date("Date", date_labels = "%Y", date_minor_breaks = "1 year")

```

Todo: Legend inside plot values, print final values, limit to 31/12/2016

## Returns analysis
TODO: Decide whether aritmetic (simple) or Logaritmic returns should be used. Here are own-defined functions which compute returns.
```{r}
retArit <- function(x) {x/lag(x)}
retLog <- function(x) {log(x)-log(lag(x))}
```

### Daily Returns Analysis
```{r, fig.width=10, fig.height=5}
perfData %>% 
  mutate_each(
    funs(retLog), c(VW=VOW, DAX=CDAX, BMW, DAI)) %>%
  select(-VOW, -CDAX) -> retDataDaily

retDataDaily %>% gather(Title, Returns, -Date) %>%
ggplot(., aes(x=Date, y=Returns, colour=Title) ) + 
  geom_line() + theme_bw() + facet_grid(~Title)
```

### Weekly Returns Analysis
```{r, fig.width=10, fig.height=5}
perfData %>% 
  arrange(Date) %>%
  mutate(week = week(Date) ) %>%
  group_by(paste0(year(Date),month(Date),week(Date))) %>%
  do(tail(., n=1)) %>%
  arrange(Date) %>% ungroup() %>%
  mutate_each(
    funs(retLog), c(VOW, CDAX, BMW, DAI)) %>%
  select(Date, VOW, CDAX, BMW, DAI) -> retDataWeekly

colnames(retDataWeekly) <- c("Date","VW","DAX","BMW","DAI")

retDataWeekly[-1, ] %>% gather(Title, Returns, -Date) %>%
  ggplot(., aes(x=Date, y=Returns, colour=Title) ) + 
  geom_line() + theme_bw() + facet_grid(~Title)
```

### Monthly Returns Analysis
```{r, fig.width=10, fig.height=5}
perfData %>% 
  arrange(Date) %>%
  mutate(month = month(Date) ) %>%
  group_by(paste0(year(Date),month(Date)) ) %>%
  do(tail(., n=1)) %>%
  arrange(Date) %>% ungroup() %>%
  mutate_each(
    funs(retLog), c(VOW, CDAX, BMW, DAI)) %>%
  select(Date, VOW, CDAX, BMW, DAI) -> retDataMonthly

colnames(retDataMonthly) <- c("Date","VW","DAX","BMW","DAI")

retDataMonthly[-1, ] %>% gather(Title, Returns, -Date) %>%
  ggplot(., aes(x=Date, y=Returns, colour=Title) ) + 
  geom_line() + theme_bw() + facet_grid(~Title)
```













